### 2022-06-15 doc/university-covid.tex
done

Mathematical models have played a crucial role in guiding pandemic responses, espeicially among higher education institutions. University campuses further present a particularly well-documented case for institutional outbreaks, thereby providing a unique opportunity to understand detailed patterns of pathogen spread.
Here, we present descriptive and modeling analyses of SARS-CoV-2 outbreaks on the Princeton University campus---this model was used throughout the pandemic to inform policy decisions for the Princeton University campus.
We demonstrate strong spatiotemporal correlations in epidemic patterns between the Princeton University campus and surrounding communities.
These findings are corroborated by our model predictions, which indicate that the amount of on-campus transmission was likely limited during much of the wider pandemic until the end of 2021.
Finally, we find that a super-spreading event played a major role in driving the recent outbreak on the Princeton University campus during the spring semester of the 2021--2022 academic year.
Despite large numbers of cases on campus, case levels in surrounding communities remained low, suggesting that there was little to no spillover transmission from campus to the local community.

@@

Predicting and controlling the spread of SARS-CoV-2 has remained a critical public health and scientific question throughout the ongoing SARS-CoV-2 pandemic \citep{baker2021limits}.
Rapid, asymptomatic transmission of SARS-CoV-2 has hindered intervention efforts, such as contact tracing \citep{hellewell2020feasibility}.
Social distancing measures have played major roles in preventing transmission but can be difficult to maintain for a prolonged period \citep{galanti2021social}.
The development of vaccines has provided a safe means of reopening society, but uncertainty remains in their long-term effectiveness in protecting infection and transmission, especially in the face of new viral variants.

Mathematical models have played a central role in guiding these pandemic responses and devising control strategies \citep{metcalf2020mathematical}.
Models can help monitor key parameters that govern epidemic dynamics \citep{kraemer2021monitoring} and retrospectively estimate the impact of intervention measures in reducing transmission \citep{flaxman2020estimating}.
These estimates can further inform projections of future scenarios and allow us to explore the endemicity of SARS-CoV-2 \citep{kissler2020projecting,saad2020immune,lavine2021immunological,saad2021epidemiological}.
% While models are useful for exploring such counterfactuals, they must be paired with real-life implementations to properly assess the effectiveness of intervention measures.

Mathematical models have also been widely deployed in planning campus reopenings.
Researchers from various institutions in the US---including Cornell \citep{frazier2022modeling}, Emory \citep{lopman2020model}, Georgia Institute of Technology \citep{gibson2021surveillance}, and UC Berkeley \citep{brook2021optimizing}---modeled the feasibility of controlling the epidemic on their campuses and considered mass asymptomatic testing as their main intervention.
These modeling efforts helped identify key parameters for control, such as the testing turnaround time, and provided support for implementing similar measures at other institutions.
Coupling modeling efforts with real-life implementations in university campuses further provided unique opportunities to directly test model-based predictions of intervention effects in preventing the transmission of SARS-CoV-2 \citep{frazier2022modeling}---
each university campus offers a relatively well-controlled epidemic setting with a behaviorally homogeneous population (especially among undergraduate students).
Mass asymptomatic testing also provides near-perfect ascertainment for epidemic sizes, allowing for accurate understanding of epidemic patterns.

University campuses present unique challenges to controlling an outbreak.
High-density environments, especially within dorms, and a large fraction of asymptomatic infections (due to the young age of university students) can readily permit rapid transmission.
These kinds of contacts are inherently difficult to keep track of, making contact tracing less effective.
The impact of intervention measures is also expected to vary across different university campuses, reflecting heterogeneity in campus settings such as compliance, resources, community prevalence, as well as effects of other interventions present on campuses.
For example, Duke and Harvard Universities experienced moderate outbreaks at the beginning of the fall semester in 2021 when in-person classes were allowed, despite high vaccination rates and weekly asymptomatic testing protocols \citep{dukeoutbreak,harvardoutbreak}, whereas the number of cases remained low in Princeton University (PU) during the same time period with similar levels of testing and vaccination.
Here, we focused on the spread of SARS-CoV-2 on PU campus alone to eliminate heterogeneities inherent to such comparisons.

In this study, we analyzed the spread of SARS-CoV-2 on PU campus (\fref{princeton}).
We first begin with a descriptive analysis of the outbreak and then present modeling analyses of the past outbreaks and future scenarios.
PU is located in Mercer County, New Jersey, USA;
the population comprises of 5267 undergraduate students, 2946 graduate students, and 7000 faculty and staff members.
For simplicity, we divided the epidemic into four time periods representing four semesters across two academic years: Fall 2020--2021 (August 24, 2020--January 1, 2021; \fref{princeton}A), Spring 2020--2021 (January 16, 2021--May 14, 2021; \fref{princeton}B), Fall 2021--2022 (August 14, 2021--December 31, 2021; \fref{princeton}C), and Spring 2021--2022 (January 1, 2022--March 18, 2022; \fref{princeton}D).
Throughout the majority of the study period, all students, faculty and staff members who were physically present for more than 8 hours on campus per week were required to participate in asymptomatic testing with varying frequencies.
Asymptomatic individuals submitted self-collected saliva samples, from which the presence of SARS-CoV-2 was tested using Reverse Transcription Polymerase Chain Reaction (RT-PCR).  
Those who tested positive were required to isolate for at least 10 days after symptom onset or test date (whichever was longer) and were released when they had been at least 48 hours with improving or resolving symptoms as per New Jersey Department of Health guidance.
PCR positives were exempt from asymptomatic testing for 90 days.
Since March 7, 2022, asymptomatic testing frequencies decreased to once a month from once a week for individuals whose vaccine statuses are up-to-date, which in turn likely reduced the accuracy of surveillance; therefore, we chose to focus on epidemic patterns before this change was implemented.
Throughout the study period, contact tracing was also performed for positive cases to alert their close contacts to either quarantine or test more frequently according to the close contacts' vaccination status, and to gather data that could help uncover clusters of transmission or superspreader events.
All data used in this analysis are publicly available on the PU COVID-19 Dashboard website: \url{https://covid.princeton.edu/dashboard}.

@@

During the fall semester of the 2020--2021 academic year, roughly 1000 grad students and 2000 faculty and staff members were present on campus and participated in asymptomatic testing. 
Undergraduate students were not allowed to return to campus as all classes were held virtually, and so their numbers were minimal on campus ($<300$).  
Both undergraduate and graduate students were required to get tested twice a week, whereas faculty and staff members were required to get tested once a week.
The number of cases remained relatively low throughout the semester with a peak occurring in early December, coinciding with the epidemic trajectory in Mercer County (\fref{princeton}A).  
A sudden decrease in the number of cases around Thanksgiving partly reflects the reduced number of tests (3852 and 2972 asymptomatic tests performed on the week ending November 20th and 27th, respectively).
The highest number of cases was reported among faculty and staff members ($=169$), followed by graduate students ($=41$), and undergraduate students ($=4$).

In the beginning of the spring semester of the 2020--2021 academic year, the number of cases suddenly increased before classes started (\fref{princeton}B), reflecting $\approx 3000$ undergraduate students returning to campus.
Returning students were required to be tested and quarantine for 7 days regardless of their returning location.
All classes remained virtual, and the testing protocol did not change (twice a week for undergraduate and graduate students, and once a week for faculty and staff members).
The number of cases persisted at similar levels as the fall semester and eventually decreased as classes ended and students went back home---the decrease in the number of cases in PU also coincided with the decrease in the number of cases in Mercer County.
The highest number of cases was reported among faculty and staff members ($=111$), followed by undergraduate students ($=101$), and graduate students ($=29$).

For the fall semester of the 2021--2022 academic year, all students and faculty and staff members were required to be vaccinated with a very few medical and religious exemptions.
By the beginning of the semesters, ??\% of students and ??\% of faculty and staff members were vaccinated \swp{ask Robin and Irini}.
% As of January 11th, 99\% of undergraduate students, 98\% of graduate students, and 96\% of faculty and staff members were vaccinated.
Vaccinees were required to be tested once a week, while unvaccinated individuals were required to be tested twice a week.
In-person classes and social events fully resumed on campus while all individuals were required to wear masks indoors with a few exceptions (e.g., when eating or drinking, or when teaching a small class).
The number of cases remained similar to previous semesters until November when the number of cases began to increase primarily among undergraduate students around Thanksgiving (\fref{princeton}C).  
In order to prevent transmission, testing frequency was increased to twice a week for undergraduate students on November 27th, 2021; the size of non-academic gatherings were also limited to 20 people.
The number of cases decreased slightly as classes ended but soon increased back up as the Omicron (BA.1) variant began to spread on campus and in Mercer county.

For the spring semester of the 2021--2022 academic year, all eligible students and faculty and staff members were required to be boosted.
By the beginning of the semesters, ??\% of students and ??\% of faculty and staff members were boosted \swp{ask Robin and Irini}.
Undergraduate students were still required to be tested twice a week to prevent the additional spread of the Omicron variant.
The number of cases remained high before classes began but decreased over time, following epidemic patterns in Mercer County (\fref{princeton}D).
As the campus and local cases numbers decreased, the gathering policy was updated to allow food in events were no longer limited to 20 people;
in addition, the testing frequency was reduced a once a week.
Following the policy change, a large outbreak occurred on campus with high case numbers persisting until the spring break.
On March 7, 2022, mask mandates were lifted and testing frequency was reduced to once a month.
The timing of this outbreak also coincided with a rapid turnover of the Omicron subvariant BA.2---the proportion of BA.2 subvariant reached ??\% compared to ??\% from the previous week \swp{Ask Dan}.

Across the first three semesters, we find strong and significant correlations between the weekly logged numbers of cases from PU and those from Mercer county:
fall 2020--2021 ($\rho = 0.81$ (95\% CI: 0.55--0.92; $p < 0.001$), \fref{princeton}E); spring 2020--2021 ($\rho = 0.80$ (95\% CI: 0.51--0.92; $p < 0.001$), \fref{princeton}F); and fall 2021--2022 ($\rho = 0.86$ (95\% CI: 0.67--0.94; $p < 0.001$), \fref{princeton}G). 
These correlations are robust even when we stratify cases by the population, except for undergraduate students during Fall 2020 when most of them were not physically present on campus (Supplementary Figure S1).
However, the case patterns in PU were decoupled from those in Mercer county for the spring semester of the 2021--2022 academic year with unclear correlations (\fref{princeton}H): $\rho = 0.32$ (95\% CI: -0.34--0.77; $p = 0.33$).
Stratifying cases by the population shows that case patterns in graduate students and faculty and staff members were still strongly correlated with case patterns in Mercer county, meaning that the campus transmission was limited to undergraduate students (Supplementary Figure S1). 
We also find strong correlations between the weekly logged numbers of cases from PU and those from other counties in New Jersey (Supplementary Figure S2)---these correlations significantly decreased with distance from Mercer county in both spring 2020--2021 ($\rho=-0.48$ (95\% CI: $-0.75$--$-0.06$; $p = 0.03$) for all cases and $\rho=-0.51$ (95\% CI: $-0.77$--$-0.10$; $p = 0.02$) for faculty and staff cases) and fall 2021--2022 ($\rho=-0.68$ (95\% CI: $-0.86$--$-0.35$, $p < 0.001$) for all cases and $\rho=-0.74$ (95\% CI: $-0.89$--$-0.46$, $p < 0.001$) for faculty and staff cases).
Across the first three semesters, both the total cases and faculty and staff cases showed similar levels of correlations with local cases. 
For the spring semester of the 2021--2022 academic year, we still find high correlations between faculty and staff cases and local cases throughout other countries ($\rho > 0.8$ across all counties in New Jersey); however, the total cases exhibit considerably weaker correlations due to student-to-student transmission on campus (Supplementary Figure S2).

@@

Finally, mass testing allows us to infer the ratio between the weekly numbers of cases from Princeton and those from Mercer county---
we expect this ratio to remain constant over time if the majority of infections on campus is caused by community transmission provided that testing patterns remain constant in both places.
We find that the ratio between the weekly numbers of cases from Princeton and those from Mercer county stayed between 1:100 and 1:25 until the end of the fall semester of the 2021--2022 academic year (\fref{princeton}I--K).
An increase in this ratio at the end of November 2021 was associated with the campus outbreak before Thanksgiving followed by an introduction of the Omicron variant in December---this deviation indicates an increase in the amount of transmission on campus.
During the spring semester of the 2021--2022 academic year, the ratio between PU cases and Mercer county cases increased above one, meaning that more cases were reported in PU than elsewhere in Mercer county (\fref{princeton}L);
notably, we did not see an increase in Mercer county cases, meaning that there was little-to-no transmission from campus to local community.

@@

We use a discrete-time, individual-based model to simulate the spread of SARS-CoV-2 on the PU campus.
This model was initially developed and used throughout the pandemic to inform policy decisions in PU, including the frequency of asymptomatic tests and the number of isolation beds required.
We continuously updated the model to reflect changes in school settings (e.g., students returning back to campus after a virtual semester) as well as intervention measures (e.g., vaccination in fall 2021 and booster shots with the emergence of the Omicron variant).
Here, we present a generic and parsimonious version that encompasses sufficient details to characterize the overall spread of SARS-CoV-2 in PU without an over-proliferation of parameters.
The model consists of four main components that are simulated on a daily time scale: (1) infection and transmission dynamics, (2) sampling and testing protocols, (3) isolation protocols, and (4) vaccination dynamics, including waning immunity and booster shots.
Previous versions of the model included contact tracing, but we exclude it in this study for simplicity.

@@

Transmission processes are modeled by first setting the contact reproduction number $\Rc$, which we define as the average number of infectious contacts an infected individual would make throughout the course of their infection;
here, infectious contacts refer to contacts that would result in infection when the contacted individual is susceptible to infection.
The contact reproduction number implicitly accounts for all intervention measures that we do not model explicitly, such as social distancing and contact tracing---in other words, this contact reproduction number does not account for asymptomatic testing or vaccination, which are modeled separately.
We further decompose $\Rc$ into pre-symptomatic $\mathcal R_p = \beta_p D_p$ and (a)symptomatic $\mathcal R_s = \beta_s D_s$ reproduction numbers, where $\beta_p$ and $\beta_s$ represent the corresponding infectious contact rates during pre-symptomatic and (a)symptomatic stages, respectively. Pre-symptomatic and (a)symptomatic reproduction numbers are calculated based on the assumed value of the proportion of presymptomatic transmission $p_p = 0.5$: $\mathcal R_p/\mathcal R_s = p_p/(1 - p_p)$.
On each day, all infected individuals who have not yet been isolated then make infectious contacts at random to anyone on campus; the number of infectious contacts are drawn from a negative binomial distribution with a mean of either $\beta_p$ or $\beta_s$ and an overdispersion parameter of $k=0.1$ to account for the possibility of super-spreading events \citep{10.12688/wellcomeopenres.15842.3}.
Infectious contacts from local community can be also made at random to anyone on campus; these contacts are modeled using a Poisson distribution with a time-varying mean, which is calculated by scaling the daily number of cases by $\theta$ and shifting it by 1 week to account for reporting delays.
Infectious contacts, whether made by individuals on campus or from outside, result in infection only when the contacted individuals are susceptible; when the contacted individuals are vaccinated, and therefore partially susceptible to infection, they have a reduced probability of infection corresponding to their susceptibility (discussed later).

All individuals on campus are assumed follow a pre-determined asymptomatic testing plan at a fixed frequency---
for example, under weekly testing, one individual can get sampled on days 1, 8, 15, and so forth, while another individual get sampled on days 2, 9, 16, and so forth.
We assume that test results come back after one day.
Symptomatic individuals can choose to take rapid PCR tests (with results returning on the same day) with a given probability on each day until their symptoms resolve---this probability is set to 1 for simulations presented in the main text.
We further assume that symptomatic individuals are isolated immediately when they submit their samples until they receive negative results.
All individuals who test positive are required to isolate (following the same isolation rule as described earlier) and are exempt from asymptomatic testing for 90 days.
Isolated individuals are assumed to no longer transmit infections.
We assume that PCR tests can detect infections from individuals who are in pre-symptomatic, symptomatic, and asymptomatic stages with 95\% sensitivity and 100\% specificity.

As most students as well as faculty and staff members had received two doses of vaccination in the beginning of fall 2021, we do not distinguish the first and second doses.
Instead, we assume that all vaccinated individuals have 90\% reduced susceptibility and 20\% reduced transmissibility at the beginning of the semester---these assumptions are consistent with recent estimates by \citep{prunas2022vaccination} that vaccination with BNT162b2 reduces susceptibility by 89.4\% (95\% CI: 88.7\%--90.0\%) and infectiousness by 23.0\% (95\% CI: $-11.3\%$--46.7\%) against the Delta strain.
Based on \citep{tartof2021effectiveness}, vaccine efficacy against susceptibility is allowed to exponentially wane from 90\% to 50\% in 20 weeks (and continues to wane at the same rate) for each vaccinated individual;
vaccine efficacy against transmissibility is also allowed to wane at the same rate (i.e., from 20\% to 11\% in 20 weeks).

In this study, we use this model to retrospectively analyze past outbreaks.
First, we try to match our model to epidemic patterns seen on campus for the first three semesters, during which there was limited campus transmission, by varying the contact reproduction number $\Rc$ and the amount of community transmission $\theta$ and holding all other parameters constant.
For each parameter combination, we simulate 100 epidemic trajectories and calculate the sum of squared differences between the weekly numbers of the observed and predicted positive cases.
The population size and testing frequencies (with twice weekly testing modeled as testing every 3 days) are set to reflect realistic campus settings.
Although we account for heterogeneity in the number of individuals in each population group on campus (i.e., undergraduate students, graduate students, and faculty and staff members) and their respective testing patterns (e.g., twice a week for undergraduate and graduate students and once a week for faculty and staff members during fall and spring, 2020), we assume that all other parameters are equal across different groups for simplicity. 
We further assume that the population mixes homogeneously.
While these assumptions are most conservative, epidemiological parameters  and mixing patterns likely differ across groups (e.g., undergraduate students are more likely to infect undergraduate students and also remain asymptomatic).
Therefore, our model parameters describe average dynamics across different groups and must be interpreted with care.

>>>>

Mathematical models have played a crucial role in guiding pandemic responses. 
University campuses present a particularly well-documented case for institutional outbreaks, thereby providing a unique opportunity to understand detailed patterns of pathogen spread.
Here, we present descriptive and modeling analyses of SARS-CoV-2 transmission on the Princeton University campus---this model was used throughout the pandemic to inform policy decisions and operational guidelines for the university campus.
We demonstrate strong spatiotemporal correlations in epidemic patterns between the university campus and surrounding communities.
These findings are corroborated by our model predictions, which indicate that the amount of on-campus transmission was likely limited during much of the wider pandemic until the end of 2021.
Finally, we find that a super-spreading event likely played a major role in driving the recent outbreak on the Princeton University campus during the spring semester of a 2021--2022 academic year.
Despite large numbers of cases on campus, case levels in surrounding communities remained low, suggesting that there was little spillover transmission from campus to the local community.

@@

Predicting and controlling the spread of SARS-CoV-2 has remained a critical public health and scientific question throughout the ongoing SARS-CoV-2 pandemic \citep{baker2021limits}.
Rapid, asymptomatic transmission of SARS-CoV-2 has hindered intervention efforts, such as contact tracing \citep{hellewell2020feasibility}.
Social distancing measures have played major roles in preventing transmission, but can be difficult to maintain for a prolonged period \citep{galanti2021social}.
The development of vaccines has provided a safe means of reopening society, but uncertainty remains in their long-term effectiveness in preventing infection and transmission, especially in the face of new viral variants.

Mathematical models have played a significant role in guiding these pandemic responses and devising control strategies \citep{cobey2020modeling,holmdahl2020wrong,metcalf2020mathematical,koelle2022changing}.
Models can help monitor key parameters that govern epidemic dynamics \citep{kraemer2021monitoring} and retrospectively estimate the impact of intervention measures in reducing transmission \citep{flaxman2020estimating}.
These estimates can further inform projections of future scenarios and allow us to explore the endemicity of SARS-CoV-2 \citep{kissler2020projecting,saad2020immune,lavine2021immunological,saad2021epidemiological}.
% While models are useful for exploring such counterfactuals, they must be paired with real-life implementations to properly assess the effectiveness of intervention measures.

Mathematical models have also been widely deployed in planning campus reopenings.
Researchers from various institutions in the US---including Cornell \citep{frazier2022modeling}, Emory \citep{lopman2020model}, Georgia Institute of Technology \citep{gibson2021surveillance}, and UC Berkeley \citep{brook2021optimizing}---modeled the feasibility of controlling the epidemic on their campuses and considered mass asymptomatic testing as their main intervention.
These modeling efforts helped identify key parameters for control, such as the testing turnaround time, and provided support for implementing similar measures at other institutions.
Coupling modeling efforts with real-life implementations in university campuses further provided unique opportunities to directly test model-based predictions of intervention effects in preventing the transmission of SARS-CoV-2 \citep{frazier2022modeling}---
each university campus offers a relatively well-controlled epidemic setting with a relatively homogeneously behaving population (especially among undergraduate students).
Campuses can also offer strong opportunities for control by non-pharmaceutical interventions, such as isolation and mask-wearing;
mass asymptomatic testing further provides robust ascertainment for epidemic sizes, allowing for accurate understanding of epidemic patterns.

On the other hand, university campuses also present unique challenges to controlling an outbreak.
A large fraction of asymptomatic infections (due to the young age of university students) and high-density interactions---such as eating in large dining halls and various social activities---can readily permit rapid transmission.
These kinds of contacts are inherently difficult to keep track of, making contact tracing less effective.
The impact of intervention measures is expected to vary across different university campuses, reflecting heterogeneity in campus settings such as compliance, resources, community prevalence, as well as effects of other interventions present on campuses.
For example, Duke and Harvard Universities experienced moderate outbreaks at the beginning of the fall semester in 2021 when in-person classes were allowed, despite high vaccination rates and weekly asymptomatic testing protocols \citep{dukeoutbreak,harvardoutbreak}, whereas the number of cases remained low in Princeton University (PU) during the same time period with similar levels of testing and vaccination.
Here, we focused on the dynamics of SARS-CoV-2 on the PU campus alone to eliminate heterogeneities inherent to such comparisons.

We analyzed the spread of SARS-CoV-2 on campus (\fref{princeton}).
We begin with a descriptive analysis of the outbreak and then present modeling analyses of the various outbreaks during 2020--2022.
PU is located in Mercer County, New Jersey, USA;
the population comprises of 5267 undergraduate students, 2946 graduate students, and around 7000 faculty and staff members.
For simplicity, we divided the epidemic into four time periods representing four semesters across two academic years: Fall 2020--2021 (August 24, 2020--January 1, 2021; \fref{princeton}A), Spring 2020--2021 (January 16, 2021--May 14, 2021; \fref{princeton}B), Fall 2021--2022 (August 14, 2021--December 31, 2021; \fref{princeton}C), and Spring 2021--2022 (January 1, 2022--March 18, 2022; \fref{princeton}D).
Throughout the majority of the study period, all students, faculty and staff members who were physically present for more than 8 hours on campus per week were required to participate in asymptomatic testing with varying frequencies.
Asymptomatic individuals submitted self-collected saliva samples, from which the presence of SARS-CoV-2 was tested using Reverse Transcription Polymerase Chain Reaction (RT-PCR).  
Those who tested positive were required to isolate for at least 10 days after symptom onset or test date (whichever was longer) and were released when they had been at least 48 hours with improving or resolving symptoms as per New Jersey Department of Health guidance.
PCR positives were exempt from asymptomatic testing for 90 days.
Since March 7, 2022, asymptomatic testing frequencies decreased to once a month from once a week for individuals whose vaccine status is up-to-date. This in turn likely reduced the accuracy of surveillance; therefore, we chose to focus on epidemic patterns before this change was implemented.
Throughout the study period, contact tracing was also performed for positive cases to alert their close contacts to either quarantine or test more frequently according to the close contacts' vaccination status, and to gather data that could help uncover clusters of transmission or superspreader events.
Changes in testing frequency and other intervention measures throughout the study period reflected various factors, including the impact of COVID-19 cases on continuity of operations or continuity of teaching; on severity of disease on campus; the capacity of testing and the healthcare system; and hospitalization rates on campus and in the area.
All data used in this analysis are publicly available on the PU COVID-19 Dashboard website: \url{https://covid.princeton.edu/dashboard}.

@@

During the fall semester of the 2020--2021 academic year, roughly 1000 grad students and 2000 faculty and staff members were present on campus and participated in asymptomatic testing. 
All classes were held virtually, and so only a few undergraduate students remained on campus ($<300$).  
Both undergraduate and graduate students were required to get tested twice a week, whereas faculty and staff members were required to get tested once a week.
The number of cases remained relatively low throughout the semester with a peak occurring in early December, coinciding with the epidemic trajectory in Mercer County (\fref{princeton}A).  
A sudden decrease in the number of cases around Thanksgiving partly reflects the reduced number of tests (3852 and 2972 asymptomatic tests performed on the week ending November 20th and 27th, respectively).
The highest number of cases was reported among faculty and staff members ($=169$), followed by graduate students ($=41$), and undergraduate students ($=4$).

In the beginning of the spring semester of the 2020--2021 academic year, the number of cases suddenly increased before classes started (\fref{princeton}B), reflecting $\approx 3000$ undergraduate students returning to campus.
Returning students were required to be tested and quarantine for 7 days regardless of their returning location.
Most classes remained virtual, and the testing protocol did not change (twice a week for undergraduate and graduate students, and once a week for faculty and staff members).
Some smaller classes were held in-person, but required social distancing (thereby limiting the size of the class) and mask-wearing at all times.
The number of cases persisted at similar levels to the fall semester and eventually decreased as classes ended and students went home---the decrease in the number of cases in PU also coincided with the decrease in the number of cases in Mercer County.
The highest number of cases was reported among faculty and staff members ($=111$), followed by undergraduate students ($=101$), and graduate students ($=29$).

For the fall semester of the 2021--2022 academic year, all students and faculty and staff members were required to be vaccinated, with very few medical and religious exemptions.
By the beginning of the semesters, 97\% of undergraduate students, 96\% of graduate students, and 94\% of faculty and staff members were vaccinated.
Vaccinees were required to be tested once a week, while unvaccinated individuals were required to be tested twice a week.
In-person classes and social events fully resumed on campus, though all individuals were required to wear masks indoors with a few exceptions (e.g., when eating or drinking, or when teaching a small class).
The number of cases remained similar to previous semesters until November when cases began to increase, primarily among undergraduate students around Thanksgiving (\fref{princeton}C).  
In order to prevent transmission, testing frequency was increased to twice a week for undergraduate students on November 27th, 2021; the size of non-academic gatherings were also limited to 20 people.
The number of cases decreased slightly as classes ended but soon increased again as the Omicron (BA.1) variant began to spread on campus and in Mercer county.

For the spring semester of the 2021--2022 academic year, all eligible students and faculty and staff members were required to be boosted.
By the beginning of the semesters, 65\% of undergraduate students, 71\% of graduate students, and 82\% of faculty and staff members were boosted.
Undergraduate students were still required to be tested twice a week to prevent the additional spread of the Omicron variant.
The number of cases remained high before classes began but decreased over time, following epidemic patterns in Mercer County (\fref{princeton}D).
Coinciding with the decrease in the campus and local cases numbers, the gathering policy was updated to allow food in events were no longer limited to 20 people;
in addition, the testing frequency was reduced to once a week.
Following the policy change, a large outbreak occurred on campus with high case numbers persisting until Spring Break (March 5th, 2022).
The timing of this outbreak also coincided with a rapid turnover of the Omicron subvariant BA.2---the proportion of BA.2 subvariant reached 93.5\% (372/398) compared to 26.9\% (14/52) from the previous week.
On March 7, 2022, mask mandates were lifted and testing frequency was reduced to once a month.

\subsection{Comparison with community transmission}

Across the first three semesters, we find strong and significant correlations between the weekly logged numbers of cases from PU and those from Mercer county:
fall 2020--2021 ($\rho = 0.81$ (95\% CI: 0.55--0.92; $p < 0.001$), \fref{princeton}E); spring 2020--2021 ($\rho = 0.80$ (95\% CI: 0.51--0.92; $p < 0.001$), \fref{princeton}F); and fall 2021--2022 ($\rho = 0.86$ (95\% CI: 0.67--0.94; $p < 0.001$), \fref{princeton}G). 
These correlations are robust even when we stratify cases by the population, except for undergraduate students during Fall 2020, when most were not physically present on campus (Supplementary Figure S1).
However, the case patterns in PU were decoupled from those in Mercer county for the spring semester of the 2021--2022 academic year with unclear correlations (\fref{princeton}H): $\rho = 0.32$ (95\% CI: $-0.34$--0.77; $p = 0.33$).
Stratifying cases by subpopulation shows that case patterns in graduate students and faculty and staff members were still strongly correlated with case patterns in Mercer county, meaning that the campus transmission was limited to undergraduate students (Supplementary Figure S1). 
We also find strong correlations between the weekly logged numbers of cases from PU and those from other counties in New Jersey (Supplementary Figure S2)---these correlations significantly decreased with distance from Mercer county in both spring 2020--2021 ($\rho=-0.48$ (95\% CI: $-0.75$--$-0.06$; $p = 0.03$) for all cases and $\rho=-0.51$ (95\% CI: $-0.77$--$-0.10$; $p = 0.02$) for faculty and staff cases) and fall 2021--2022 ($\rho=-0.68$ (95\% CI: $-0.86$--$-0.35$, $p < 0.001$) for all cases and $\rho=-0.74$ (95\% CI: $-0.89$--$-0.46$, $p < 0.001$) for faculty and staff cases).
Across the first three semesters, both the total cases and faculty and staff cases showed similar levels of correlations with local cases. 
For the spring semester of the 2021--2022 academic year, we still find high correlations between faculty and staff cases and local cases throughout other countries ($\rho > 0.8$ across all counties in New Jersey); however, the total cases exhibit considerably weaker correlations due to student-to-student transmission on campus (Supplementary Figure S2).

@@

Finally, mass testing allows us to infer the ratio between the weekly numbers of cases from Princeton and those from Mercer county---
we expect this ratio to remain constant over time if the majority of infections on campus is caused by community transmission, provided that testing patterns remain roughly constant in both places.
We find that the ratio between the weekly numbers of cases from Princeton and those from Mercer county stayed between 1:100 and 1:25 until the end of the fall semester of the 2021--2022 academic year (\fref{princeton}I--K).
An increase in this ratio at the end of November 2021 was associated with the campus outbreak before Thanksgiving followed by an introduction of the Omicron variant in December---this deviation indicates an increase in the amount of transmission on campus.
During the spring semester of the 2021--2022 academic year, the ratio between PU cases and Mercer county cases increased above one, meaning that more cases were reported in PU than elsewhere in Mercer county (\fref{princeton}L);
notably, we did not see an increase in Mercer county cases, meaning that there was little-to-no transmission from campus to local community.

@@

We use a discrete-time, individual-based model to simulate the spread of SARS-CoV-2 on the PU campus.
This model was initially developed and used throughout the pandemic to inform policy decisions in PU, including the frequency of asymptomatic tests and the number of isolation beds required.
We continuously updated the model to reflect changes in school settings (e.g., students returning back to campus after a virtual semester) as well as intervention measures (e.g., vaccination in fall 2021 and booster shots with the emergence of the Omicron variant).
Here, we present a generic and parsimonious version that encompasses sufficient details to characterize the overall spread of SARS-CoV-2 in PU without an over-proliferation of parameters.
The model consists of four main components simulated on a daily time scale: (1) infection and transmission dynamics, (2) sampling and testing protocols, (3) isolation protocols, and (4) vaccination dynamics, including waning immunity and booster shots.
Previous versions of the model included contact tracing, but we exclude it in this model for simplicity.

@@

Transmission processes are modeled by first setting the contact reproduction number $\Rc$, which we define as the average number of infectious contacts an infected individual would make throughout the course of their infection;
here, infectious contacts refer to contacts that would result in infection when the contacted individual is susceptible to infection.
The contact reproduction number implicitly accounts for all intervention measures that we do not model explicitly, such as social distancing and contact tracing---in other words, this contact reproduction number does not account for asymptomatic testing or vaccination, which are modeled separately.
We further decompose $\Rc$ into pre-symptomatic $\mathcal R_p = \beta_p D_p$ and (a)symptomatic $\mathcal R_s = \beta_s D_s$ reproduction numbers, where $\beta_p$ and $\beta_s$ represent the corresponding infectious contact rates during pre-symptomatic and (a)symptomatic stages, respectively. Pre-symptomatic and (a)symptomatic reproduction numbers are calculated based on the assumed value of the proportion of presymptomatic transmission $p_p = 0.5$: $\mathcal R_p/\mathcal R_s = p_p/(1 - p_p)$.
On each day, all infected individuals who have not yet been isolated then make infectious contacts at random to anyone on campus; the number of infectious contacts are drawn from a negative binomial distribution with a mean of either $\beta_p$ or $\beta_s$ and an overdispersion parameter of $k=0.1$ to account for the possibility of super-spreading events \citep{10.12688/wellcomeopenres.15842.3}.
We also rely on cases from Mercer County to crudely capture community dynamics.
In particular, we assume that infectious contacts from local or regional community can be made at random to anyone on campus; these contacts are modeled using a Poisson distribution with a time-varying mean, which is calculated by scaling the daily number of cases by $\theta$ and shifting it by 1 week to account for reporting delays.
Infectious contacts, whether made by individuals on campus or from outside, result in infection only when the contacted individuals are susceptible; when the contacted individuals are vaccinated, and therefore partially susceptible to infection, they have a reduced probability of infection corresponding to their susceptibility (discussed later).

All individuals on campus are assumed to follow a pre-determined asymptomatic testing plan at a fixed frequency---
for example, under weekly testing, one individual can get sampled on days 1, 8, 15, and so forth, while another individual get sampled on days 2, 9, 16, and so forth.
We assume that test results come back after one day.
Symptomatic individuals can choose to take rapid PCR tests (with results returning on the same day) with a given probability on each day until their symptoms resolve---this probability is set to 1 for simulations presented in the main text.
We further assume that symptomatic individuals are isolated immediately when they submit their samples until they receive negative results.
All individuals who test positive are required to isolate (following the same isolation rule as described earlier) and are exempt from asymptomatic testing for 90 days.
Isolated individuals are assumed to no longer transmit infections.
We assume that PCR tests can detect infections from individuals who are in pre-symptomatic, symptomatic, and asymptomatic stages with 95\% sensitivity and 100\% specificity.

As most students, as well as faculty and staff members, had received two doses of vaccination in the beginning of fall 2021, we do not distinguish the first and second doses.
Instead, we assume that all vaccinated individuals have 90\% reduced susceptibility and 20\% reduced transmissibility at the beginning of the semester---these assumptions are consistent with recent estimates by \cite{prunas2022vaccination} that vaccination with BNT162b2 reduces susceptibility by 89.4\% (95\% CI: 88.7\%--90.0\%) and infectiousness by 23.0\% (95\% CI: $-11.3\%$--46.7\%) against the Delta strain.
Based on \cite{tartof2021effectiveness}, vaccine efficacy against susceptibility is allowed to exponentially wane from 90\% to 50\% in 20 weeks (and continues to wane at the same rate) for each vaccinated individual;
vaccine efficacy against transmissibility is also allowed to wane at the same rate (i.e., from 20\% to 11\% in 20 weeks).

In this study, we use this model to retrospectively analyze past outbreaks.
First, we try to match our model to epidemic patterns seen on campus for the first three semesters, during which there was limited campus transmission, by varying the contact reproduction number $\Rc$ and the amount of community transmission $\theta$ and holding all other parameters constant.
For each parameter combination, we simulate 100 epidemic trajectories and calculate the sum of squared differences between the weekly numbers of the observed and predicted positive cases.
The population size and testing frequencies (with twice weekly testing modeled as testing every 3 days) are set to reflect realistic campus settings.
Although we account for heterogeneity in the number of individuals in each population group on campus (i.e., undergraduate students, graduate students, and faculty and staff members) and their respective testing patterns (e.g., twice a week for undergraduate and graduate students and once a week for faculty and staff members during fall and spring, 2020), we assume that all other parameters are equal across different groups for simplicity. 
We further assume that the population mixes homogeneously.
While these assumptions are most parsimonious, epidemiological parameters and mixing patterns likely differ across groups (e.g., undergraduate students are more likely to infect undergraduate students and also remain asymptomatic).
Therefore, our model parameters describe average dynamics across different groups and must be interpreted with care.
